---
title: "03_analyzing_respirometry_data.Rmd"
author: "Aidan Coyle"
date: "6/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script is meant to analyze respirometry data previously collected as part of Experiment 2. 

Respiration data was collected on 10 experimental mussels from each experimental group at the start of the experiment (1-20) and all living experimental mussels, all control mussels, and some un at the end of the experiment. 


First, load all required libraries, installing those that aren't yet loaded

```{r libraries, message = FALSE, warning=FALSE}
## clear workspace
rm(list=ls())

# Add all required libraries that are installed with install.packages() here
load.lib <- c("tidyverse", "readxl","bestNormalize","agricolae","multcomp")
install.lib <- load.lib[!load.lib %in% installed.packages()] # Select missing packages
for(lib in install.lib) install.packages(lib,dependencies=TRUE) # Install missing packages + dependencies
sapply(load.lib,require,character=TRUE) # Load all packages.

## Set ggplot theme
my_theme <- theme(line              = element_line(size=1.2),
                  rect              = element_rect(size=1.2),
                  text              = element_text(size=14,color="black"),
                  panel.background  = element_blank(),
                  panel.grid.major  = element_blank(), 
                  panel.grid.minor  = element_blank(),
                  axis.text.x       = element_text(size=16,color="black"), #,angle=90),
                  axis.text.y       = element_text(size=16,color="black"),
                  axis.title.x      = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)),
                  axis.title.y      = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)),
                  axis.ticks.x      = element_line(color="black"),
                  axis.ticks.y      = element_line(color="black"),
                  # axis.line         = element_line(color = "black", size = 0.1),
                  panel.border      = element_rect(color = "black", fill=NA, size=1.2)
                  # legend.key        = element_blank(), # removes background of legend bullets
                  # legend.position   = "none"
                  ) 
```
# Read in data

```{r}
# Read in respiration data
resp_dat <- read_excel("output/EXP_2/respirometry/respirometry_summary.xlsx")

# Change date from linear to binary
resp_dat$date <- as.character(resp_dat$date)
```

# Graph 1: Boxplots of time and treatment
```{r}
### boxplots, controls
bp1 <-  ggplot(resp_dat, aes(x=as.factor(date), y=umol_L_hr_cm, 
                                       group=as.factor(bp_list), fill=bp_list)) +
        geom_boxplot(colour = "black", size = 0.8,outlier.colour="black", outlier.shape = 16,
                     outlier.size=1, notch=FALSE, varwidth = TRUE, alpha = 0.9) +
        scale_fill_manual(values=c("orangered1","red","grey","orange")) +
        # scale_y_continuous(breaks = seq(0, 2.5, 0.5), limits = c(0, 2.7)) +
        # scale_x_continuous(breaks = seq(0, 30, 5), limits = c(0, 32)) + 
        xlab(NULL) + ylab(NULL) + 
        my_theme

bp1

# ggsave("output/EXP_2/respirometry/bp_resp.png",
#        plot   = bp1,
#        dpi    = 1200,
#        device = "png",
#        width  = 4,
#        height = 4,
#        units  = "in")

```
# Graph 2: Scatter plot of value before and after treatment
```{r}
library(tidyverse)
#install.packages("gapminder")
library(gapminder)
resp_dat %>% 
  ggplot(aes(date, umol_L_hr_cm))+
  geom_point(aes(color=trt), size=3)+
  geom_line(aes(group=mussel_ID), color="grey")+theme_bw()

```

# Statistical Testing
```{r}
# Define dataset for analysis, factors
dat_stat <- resp_dat
test_me <- dat_stat$umol_L_hr_cm

dat_stat$ID         <- factor(dat_stat$mussel_ID)
dat_stat$trt        <- factor(dat_stat$trt_stat)
dat_stat$timepoint  <- factor(dat_stat$date)

# Test for normality
qqnorm(test_me) # check linearity of transformed data
qqline(test_me)
norm_test <- shapiro.test(test_me) # p-value fail = good, don't need transformation
print(norm_test$p.value)

# Normalize data if normality test failed
if(norm_test$p.value<0.05)     {
        normalized <- bestNormalize(test_me)
        test_me <- normalized$x.t # overwrite
        qqnorm(test_me) # check linearity of transformed data
        qqline(test_me)
        print(shapiro.test(test_me))
        print("transformed!",quote=FALSE)}

dat_stat$response <- test_me

# Run repeated measures anova using oyster ID has random effect
# https://www.r-bloggers.com/2015/08/two-way-anova-with-repeated-measures/
my_test <- aov(response ~ timepoint * trt + Error(1/ID),
                       data = dat_stat)

my_test_summary <- summary(my_test)
summary(my_test)

# # Significant interaction observed between timepoint:trt:ploidy. Proceed with HSD test.
# tx <- with(dat_stat, interaction(timepoint,trt,ploidy))
# amod <- aov(response ~ tx, data = dat_stat)
# mult_comp <- HSD.test(amod, "tx", group=TRUE, console=TRUE)

# # save output
# write.table(my_test_summary$`Error: Within`[[1]], file = "stats/respirometry/AOV_respirometry.csv",      row.names = TRUE)
# write.table(mult_comp$groups,                     file = "stats/respirometry/HSD_test_respirometry.csv", row.names = TRUE)


```




