---
title: "03_analyzing_respirometry_data.Rmd"
author: "Aidan Coyle"
date: "6/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script is meant to analyze respirometry data previously collected as part of Experiment 2. 

Respiration data was collected on 10 experimental mussels from each experimental group at the start of the experiment (1-20) and all living experimental mussels, all control mussels, and some un at the end of the experiment. 


First, load all required libraries, installing those that aren't yet loaded

```{r libraries, message = FALSE, warning=FALSE}
# Add all required libraries that are installed with install.packages() here
list.of.packages <- c("tidyverse", "readxl")
# Add all libraries that are installed using BiocManager here
bioconductor.packages <- ""

# Install BiocManager if needed
if(!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")

# Get names of all required packages that aren't installed
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[, "Package"])]
new.bioc.packages <- bioconductor.packages[!(bioconductor.packages %in% installed.packages()[, "Package"])]
# Install all new packages
if(length(new.packages)) install.packages(new.packages)
if(length(new.bioc.packages)) BiocManager::install(new.bioc.packages)

# Load all required libraries
all.packages <- c(list.of.packages, bioconductor.packages)
lapply(all.packages, FUN = function(X) {
  do.call("require", list(X))
})

```
# Read in data

```{r}
# Read in respiration data

resp_dat <- read_excel("../../output/EXP_2/respirometry/respirometry_summary.xlsx")

# Change date from linear to binary
resp_dat$date <- as.character(resp_dat$date)
```

# Graph 1: Bar graph of time and treatment


```{r}
library(ggplot2)
pdf(file = "../../output/EXP_2/respirometry/results_bar_graph.pdf")
ggplot(data = resp_dat, aes(x = date, y = umol_L_hr_cm, fill = trt)) +
  geom_bar(position = "dodge", stat = "identity")

dev.off()
```


# Graph 2: Scatter plot of value before and after treatment

```{r}
library(tidyverse)
#install.packages("gapminder")
library(gapminder)
resp_dat %>% 
  ggplot(aes(date, umol_L_hr_cm))+
  geom_point(aes(color=trt), size=3)+
  geom_line(aes(group=mussel_ID), color="grey")+theme_bw()

```

