---
title: "Protocol for Producing cDNA from RNA"
author: "Sarah Tanja and Aidan Coyle"
date: "5/31/2022"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tinytex)
```


## Background

For each RNA sample, it is recommended to produce cDNA in duplicate or triplicate 
to guard against potential contamination or pipetting errors. 

### Supplies Needed
- RNA samples (keep on ice until using)
- Sigma water
- oligo dT primer
- Thermal cycler
- MMLV buffer
- dNTP 
- MMLV reverse transcriptase

# Protocol

### Step 1: Determine your dilution

When preparing your sample, aim for 100$ng$ of RNA in 18 $\mu l$ of $H_{2}O$.
If encountering errors, can use the specific RNA concentration for the sample.
However generally, using the average RNA concentration for all samples prepped is close enough.

Ex: If qubit scores average 100 $\frac{ng}{\mu l}$, within each tube you will add 1 $\mu$l of RNA to 17 $\mu$l of sigma $H_{2}O$. 


If qubit scores average 50 $\frac{ng}{\mu l}$, within each tube you will add 2 $\mu$l of RNA 
to 16 $\mu$l of sigma $H_{2}O$

Or, to describe it as a formula, 

$$ RNA (\mu l) = \frac{100 ng }{qubit score (\frac{ng}{\mu l}) } $$
$$ sigma H_{2}O (\mu l) = 18 \mu l - \frac{100 (ng) }{qubit score (\frac{ng}{\mu l})} -  $$



### Step 2: Dilute, Add Primer, Incubate
- [ ] Label 0.5ml tubes with RNA sample ID (ex. 001) and a duplicate indicator (ex. a, b, c)
- [ ] Add sigma water to each 0.5ml tube (quantity determined as above - if RNA is at 50 $\frac{ng}{\mu l}$, add 16 $\mu$l sigma water)
- [ ] Add a quantity of RNA sample to its corresponding 0.5ml tubes (quantity determined as above, if RNA is at 50 $\frac{ng}{\mu l}$, add 2$\mu$l sample )
- [ ] Add 1 $\mu$l of oligo dT primer to each 0.5ml tube
- [ ] Incubate at 70$^{\circ}$C for 5 minutes

### Step 3: Create cDNA
Your master mix will consist of the following components, numbered by the number of sample vials you are working with (plus a few extra so you don't run out of master mix).


- 5 $\mu$l MMLV buffer

- 1.25 $\mu$l dNTP

- 0.5 $\mu$l MMLV reverse transcriptase

Ex: If you had 18 samples and wanted to prepare master mix for 20, your master mix would be 100 $\mu$l MMLV buffer, 25 $\mu$l dNTP, and 10 $\mu$l MMLV


- [ ] Calculate the amounts required for your master mix
- [ ] Create your master mix in a tube
- [ ] Add 6.75 $\mu$l of master mix to each sample vial
- [ ] Incubate at 42$^{\circ}$C for 1 hour
- [ ] Incubate at 95$^{\circ}$C for 3 minutes 
- [ ] Store frozen at -20$^{\circ}$C or below

# Method Narrative

We took 11 extracted gill-tissue RNA samples from each of treatments A (chronic 18C) and B (variable cycling) and 10 extracted RNA samples from the control group C ()

# Detailed Narrative

### Batch 1
Date: 05/31/22

Lab work: Aidan Coyle and Sarah Tanja

Number of samples completed: 18

Duplicates per sample: 2

Summary: Completed 18 samples in duplicate,and labeled the duplicates a & b
Not all samples had qubit scores, so we used the qubit scores generated by Chris Mantegna as a subsample. 
They ranged from 4.8 $\frac{ng}{\mu l}$ to 200 $\frac{ng}{\mu l}$ and averaged 52 $\frac{ng}{\mu l}$ of RNA. To get 100ng of RNA, we added 2$\mu$l of RNA extract to 16$\mu$l of sigma water. 

**Due to an error, added 2$\mu$l of oligo dT primer instead of 1$\mu$l!**

### Batch 2
Date: 06/01/22

Lab work: Sarah Tanja

Number of samples completed: 14

Duplicates per sample: 2

**Incubated for 10min instead of 5 at 70$^{\circ}$C :(**

Master Mix:
5$\mu$l buffer * 30 = 150$\mu$l
1.25$\mu$l dNTP * 28 = 37.5$\mu$l
0.5$\mu$l MMLV *28 = 15$\mu$l


# Equipment
All incubation was performed with a Peltier Thermal Cycler (PTC-200)

Samples were stored in at -20$^{\circ}$C